CudeleFS: Programmable Consistency and \\Durability in a Global Namespace

### 2017-04-04 paper/sections/3_posix_overheads.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the metadata servers. Each metadata server maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Journal Segment Size} \label{fig:slowdown-journal-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
      \caption{180MB Segment vs. No Journal}
      \label{fig:slowdown-journal-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
      \caption{Inode Cache}
      \label{fig:slowdown-journal-c}
  \end{subfigure}
  \caption{The overhead of global durability in CephFS. The segment size
  is the threshold that the metadata server starts
  trimming.\label{fig:slowdown-journal}}
\end{figure*}

% Effects on performance
Figure~\ref{fig:slowdown-journal} shows that journaling metadata updates into
the object store has an overhead. Figure~\ref{fig:slowdown-journal-a} shows the
effect of journaling of different journal segment sizes; the larger the segment
size the bigger that the writes into the object store are. The trade-off for
better performance is memory consumption because larger segments take up
more space for buffering. Figure~\ref{fig:slowdown-journal-b} shows how
the metadata server periodically stops serving requests to flush ({\it i.e.}
apply journal updates to) to the metadata store.  The journal overhead is
sufficient enough to slow down metadata throughput but not so much as to
overwhelm the bandwidth of the object store. We measured our peak bandwidth to
be 100MB/s, which is the speed of our network link.\\

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
      \caption{Inteference forces \texttt{lookup()}s}
      \label{fig:interfere-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
      \caption{Interference hurts performance}
      \label{fig:interfere-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-types.png}
      \caption{Different interferring requests}
      \label{fig:interfere-c}
  \end{subfigure}
  \caption{The overhead of strong consistency in CephFS. When a second client
  ``interferes", capabilities are revoked and metadata servers do more work.
  \label{fig:interfere}}
\end{figure*}

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:interfere-a}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but the overall performance decreases.  This degradation is shown in
Figure~\ref{fig:interfere-b}, where we scale the number of clients and show
increased runtime and variability. The performance is compared to a single
isolated client and the error bar is the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit. Zooming in on runs
with 7 clients in Figure~\ref{fig:interfere-c} we see how different
interfering operations affect performance, again when compared to the
performance of a single isolate client.  ``create" has little effect but
``stat" and ``create many" cause noticeable slowdowns because of the extra work
they impose on the metadata server.\\

>>>>

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
      \caption{Interference forces \texttt{lookup()}s}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
      \caption{Interference hurts variability}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{The overhead of global durability and strong consistency in CephFS.
  (a) shows the effect of different journal segment sizes, which are streamed
  into the object store for fault tolerance. (b) and (c) show that when a second
  client ``interferes", capabilities are revoked and metadata servers do more
  work.  \label{fig:overhead}}
\end{figure*}

@@

\begin{figure}[tb] \centering
\includegraphics[width=1\linewidth]{./figures/journal.png} 
\caption{CephFS has two views of the file system namespace: a journal of
metadata updates and a metadata store. For fault tolerance, they are stored in
the object store as segments and fragments, respectively.
\label{fig:journal}}
\end{figure}

% Effects on performance
Figure~\ref{fig:overhead-a} shows the
effect of journaling of different journal segment sizes; the larger the segment
size the bigger that the writes into the object store are. The trade-off for
better performance is memory consumption because larger segments take up
more space for buffering. When journaling is on, 
the metadata server periodically stops serving requests to flush ({\it i.e.}
apply journal updates) to the metadata store.  The journal overhead is
sufficient enough to slow down metadata throughput but not so much as to
overwhelm the bandwidth of the object store. We measured our peak bandwidth to
be 100MB/s, which is the speed of our network link.\\

@@

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but the overall performance decreases.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
increased variability. The performance is compared to a single
isolated client and the \(y\) axis the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit.\\

### 2017-04-04 paper/sections/4_methodology.tex
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Create            & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{Cudele's mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

@@

\subsection{Cudele's Mechanisms}
\label{sec:cudeles-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Create"
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

@@

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Create" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

@@

\subsection{Defining Policies in Cudele}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}

@@

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{creates+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

@@

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

@@

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "create+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

>>>>

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{The CudeleFS mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

@@

\subsection{The CudeleFS Mechanisms}
\label{sec:the-cudelesfs-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
description of what each mechanism does.  Decoupled clients use the ``Append Client Journal""
mechanism to append metadata updates to a local, in-memory journal. Once the
job is complete, the system calls CudeleFS mechanisms to achieve the desired
consistency/durability semantics.  CudeleFS provides a library for clients to
link into and all operations are performed by the client.

@@

%The metadata store and journal
%are different ways of representing the namespace.  CudeleFS presents 6
%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
%Persist. ``RPCs" does round trip remote procedure calls to establish
%consistency; it is the default implementation for complying with POSIX in
%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
%into the object store. ``Append Client Journal"" allows clients to append metadata events to an
%in-memory journal. ``Volatile apply" 
% ``Local Persist" takes the in-memory journal and writes it to the
%client's disk. ``Global Persist" saves the journal as a an object in the object
%store from the client.  Next, we discuss how these mechanisms can be composed
%to get different consistency and durability semantics.

@@

\subsection{Defining Policies in CudeleFS}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
  none                       & append client journal            & append client journal          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & append client journal            & append client journal          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & append client journal            & append client journal          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}

@@

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

@@

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "append_client_journal"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

@@

\begin{listing}[h]
\begin{minted}[frame=single,
               framesep=3mm,
               xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "append_client_journal+volatile_apply"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

### 2017-04-04 paper/sections/5_implementation.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Create",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemeted by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, CudeleFS leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\noindent\textbf{CudeleFS}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Create", ``Volatile Apply", and ``Nonvolatile Apply".

>>>>

\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Append Client Journal"",
``Local Persist", and ``Global Persist".  ``Stream" is the default
implementation for achieving global durability in CephFS but the rest of the
mechanisms are implemeted by writing with the journal format.  By writing with
the same format, the metadata servers can read and use the recovery code to
materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
These implementations required no changes to CephFS because the metadata
servers know how to read the events the library is writing.  By re-using the
journal subsystem to implement the namespace decoupling, CudeleFS leverages the
write/read optimized data structures, the formats for persisting events
(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
for replaying events onto the internal namespace data structures.

@@

\noindent\textbf{CudeleFS}: the external library the clients link into is based
on the journal tool. It already had functions for importing, exporting, and
modifying the updates in the journal so we re-purposed that code to implement
``Append Client Journal"", ``Volatile Apply", and ``Nonvolatile Apply".

### 2017-04-04 paper/sections/6_results.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
The performance of the CudeleFS mechanisms normalized to the runtime of the
create mechanism. The runtime of the create mechanism is the time it takes to
write 100 file creates to the client's in-memory journal of metadata
updates.\label{fig:slowdown-mechanisms}}
\end{figure}

@@

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
mechanisms, normalized to the time it takes to write 100K file create updates
to the client's in-memory journal ({\it i.e.} the ``Create" mechanism). Bars
above \(10^0\) are slower than the ``Create" mechanism and bars below are
faster.  ``Stream" is an approximation of the overhead and is calculated by
subtracting the runtime of the job with the journal turned off from the runtime
with the journal turned on. All the slowdowns and speedups reported are for the
100K file create job, the largest workoad we tested\footnote{We stopped testing
at 100K because this is the maximum recommended size of a directory in CephFS;
preliminary experiemnts with larger directory sizes show memory problems.}.

@@

Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
weak and strong consistency, normalized to the runtime of the ``Create"
mechanism (again, just creating files in the client's in-memory journal).  We
use the following compositions from the mechanisms in
Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Creates" +
``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Creates" +
``Local Persist" + ``Volatile Apply".

>>>>

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
The performance of the CudeleFS mechanisms normalized to the runtime of the
``Append Client Journal" mechanism (the runtime of writing \(n\) file creates
to the client's in-memory journal).  \label{fig:slowdown-mechanisms}}
\end{figure}

@@

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
mechanisms, normalized to the time it takes to write \(n\) file create updates
to the client's in-memory journal ({\it i.e.} the ``Append Client Journal""
mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal""
mechanism and bars below are faster.  ``Stream" is an approximation of the
overhead and is calculated by subtracting the runtime of the job with the
journal turned off from the runtime with the journal turned on.  The largest
workoad we tested is 100K file creates in the same directory.  This is the
maximum recommended size of a directory in CephFS; preliminary experiemnts with
larger directory sizes show memory problems.

\subsubsection{Poorly Scaling Data Structures} Despite doing the same amount of
work, mechanisms that rely on poorly scaling data structures have large
slowdowns for the larger number of creates. For example, ``RPCs" which relies
on the internal CephFS directory structures, goes from a \(40\times\) slowdown
for 10K files to a \(90\times\) slowdown for 100K files. It is a well-known
problem that directory data structures do not scale when creating files in the
same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
structures will experience similar slowdowns. Other mechanisms that write
events to a journal ({\it e.g.} the persists, ``Volatile Apply") experience a
much less drastic slowdown because the journal data structure does not need to
be scanned for every operation.

@@

Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
weak and strong consistency, normalized to the runtime of the ``Append Client Journal""
mechanism (again, just creating files in the client's in-memory journal).  We
use the following compositions from the mechanisms in
Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal"s" +
``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal"s" +
``Local Persist" + ``Volatile Apply".
